---
title: "Homework 1"
author: "Malika Top"
date: "2024-09-18"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
```

## Problem 1
```{r peng_explo}
data("penguins", package = "palmerpenguins")
mean_flipper_length <- mean(pull(penguins, flipper_length_mm), na.rm = TRUE)
```

The `penguins` dataset is a dataframe containing information about
adult foraging penguins near  Palmer Station, Antarctica. Some important
variables are `r colnames(penguins)`. `species` is a factor variable taking on 3
values, denoting the penguin species, Adélie, Chinstrap and Gentoo, `island` tells
us which island it came from (Biscoe, Dream, or Torgersen), and various others 
are continuous, quantitative variables characterizing the penguin's body. There are
`r nrow(penguins)` rows and `r ncol(penguins)` columns. The mean flipper length is
`r mean_flipper_length`. 

Here is a scatterplot of the penguins' bill length against their flipper length,
color-coded by their species. 
```{r peng_plot, warning=FALSE}
peng_plot <- ggplot(penguins, aes(x=bill_length_mm, y=flipper_length_mm,
                     color = species))+geom_point()
peng_plot
ggsave(peng_plot, file="peng_plot.png")
```


## Problem 2
```{r df_create}
set.seed(123)
df <- tibble(rand_sample=rnorm(10, mean=0, sd=1), 
             vec_logical = rand_sample < 0,
             vec_char = c("Hi", "my", "name", "is", "Malika", 
                          "and", "I", "am", "studying", "biostatistics"),
            vec_factor = factor(c("rice", "noodle", "potato",
                                  "potato", "rice", "noodle",
                                  "potato", "rice", "rice", 
                                  "noodle")))
```
If we try to take the mean of each variable in this dataframe, we will encounter
errors for `vec_char` and `vec_factor` because they are not of type numeric
or logical. R only calculates the mean for these two types of variables, which
in this case is vec_logical (even though it is just trues and falses, R 
automatically calculates True as 1 and False as 0) with mean 
`r mean(pull(df, vec_logical))`
and rand_sample with mean `r mean(pull(df, rand_sample))`. 

```{r df_redo, eval=FALSE}
num_logical = as.numeric(pull(df,vec_logical))
mean(num_logical)
num_char = as.numeric(pull(df, vec_char))
mean(num_char)
num_factor = as.numeric(pull(df, vec_factor))
mean(num_factor)
```
If we converted the variables of type logical, character, and factor, to be numeric
using the `as.numeric` function, we would observe the same answer for the first 
type but different answers for the second two. Converting `vec_logical` to be
numeric means that the new variable has now switched from Trues and Falses to 
1s and 0s, respectively, but the mean is still the same (`r mean(as.numeric(pull(df,vec_logical)))`)
as explained in the previous part. Converting `vec_character` to numeric doesn't 
give us an error,
but a warning message because the values just become NA as there is no way for
R to make these text characters (that contain no numbers) into a number. This would
be different if some of the values in `vec_character` included integers or real
numbers. Converting `vec_factor` to numeric does get us an actual mean though, of
`r mean(as.numeric(pull(df, vec_factor)))`. This is because a factor variable is encoded with numeric
codes, and when converted, we obtain the corresponding numeric code.

